%!TEX encoding = UTF-8 Unicode

%% Thesis cover pages
%%
%% Based on the example.tex file provided by Tomás Oliveira e Silva
%%
%% Modified by João Paulo Barraca <jpbarraca@ua.pt>
%% Modified by Filipe Manco <filipe.manco@ua.pt>

% Remove 'coverdepname' parameter to remove department name from cover pages
% Add 'final' parameter to remove "DOCUMENTO PROVISÓRIO" from cover page
\documentclass[11pt,twoside,a4paper,final]{report}

%%% This file is a template for the cover of your
%%%  thesis.
%%% Lines like this one, starting with %%% , are
%%%  instructions. Search them to know where and
%%%  how to edit this file. After editing the
%%%  document you can delete them.
%%% Items marked with OPTIONAL can be removed.
%%% Most of the lines you will have to change
%%%  already have examples to help you understand.
%%% The complete list of options and commands can
%%%  be found in uaThesisTemplate.sty


%%% Select your area of study and department from
%%%  the following lists and replace them in the
%%%  \usepackage declaration ([area,department]).
%%%
%%% | area        |   | department |
%%% ---------------   --------------
%%% | arts        |   | dao        |
%%% | sciences    |   | dbio       |
%%% | education   |   | dcspt      |
%%% | economy     |   | deca       |
%%% | engineering |   | degeit     |
%%% | humanities  |   | de         |
%%% | health      |   | deti       |
%%% |             |   | demac      |
%%% |             |   | decivil    |
%%% |             |   | dem        |
%%% |             |   | dfis        |
%%% |             |   | dlc        |
%%% |             |   | dmat       |
%%% |             |   | dq         |
%%% |             |   | dcm        |
%%% |             |   | dgeo        |
%%% ---------------   --------------
%% Import the UA thesis template
\usepackage[engineering,deti,final]{uaThesisTemplate}


%% Optional packages for utf8 encoding and T1 fonts
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

%% For prettier pages
\usepackage{lmodern}

%% We will use english and portuguese
\usepackage[english,portuguese]{babel}

\usepackage{xspace}% used by \sigla

%%% This package is only used to create a dummy
%%%  text for the abstract. You can remove the
%%%  following line in the final document.
%\usepackage{blindtext}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       DEFINE THESIS COVER FIELDS
%%                  START
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Edit thesis year
%% Thesis Year
\ThesisYear{2024}


%%% Uncomment one of the following options:
% \DocumentTypeThesis
% \DocumentTypeThesisProposal
\DocumentTypeDissertation

%%% Uncomment if you are a student of a MAP Doctoral program 
%%% http://www.map.edu.pt/
%%% | area                | identifier |
%%% | Biology of Plants   | bioplant   |
%%% | Computer Science    | i          |
%%% | Physics             | fis        |
%%% | Telecommunications  | tele       |

%\MAPProgram{i}

%%% Edit author's name
%% Author's full name
\Author{Jo\~ao Luís Fernandes Clemente}

%%% Edit thesis title in Portuguese
%% Thesis Mandatory Title in Portuguese
\TitlePT{Uso de um braço robótico para auxiliar cenários de Colaboração Remota
apoiada por Realidade Mista}

%%% OPTIONAL: you can write your thesis title in
%%%  other languages. You can repeat the following
%%%  line as many times as you want, or remove it
%%%  if you only want the portuguese title.
%% Optional Titles in additional languages.
\TitleEN{Using a Robotic Arm to Assist during Scenarios of Remote
Collaboration supported by Mixed Reality}

%%% Edit your supervisor's information.
%%%
%%% Use this command for male supervisor
%%% \Supervisor{name}{category}
%%%
%%% Or this command for female supervisor
%%% \Supervisora{name}{category}
%%%
%%% name   : Supervisor's name
%%% categoy: One of the following list
%%%
%%% | category                  |
%%% -----------------------------
%%% | catedrático/a             |
%%% | catedrático/a convidado/a |
%%% | associado/a c/ agregação  |
%%% | associado/a               |
%%% | auxiliar                  |
%%% | auxiliar convidado        |
%%% -----------------------------
%% Supervisor information
\Supervisor{Bernardo Marques}{auxiliar}

%%% OPTIONAL: you can add a second supervisor.
%%%
%%% Use this command for male supervisor
% \SupervisorMember{Eurico Pedrosa}{auxiliar}

% \SupervisorMember{name}{category}
%%% 
%%% Or this command for female supervisor
%%% \SupervisoraMember{name}{category}
%%%  {[preposition: (at, in, no, da)] department}
%%%  {[preposition: (at, in, do, na)] school}
%%%  {[preposition: (and, e)] }
%%%
%%% name      : See above.
%%% category  : See above.
%%% department: Can be a department, faculty or
%%%             other; Choose the preposition
%%%             accordingly.
%%% school    : University/Scholl name; Choose the
%%%             preposition accordingly.
%%% last      : Write {e } (Portuguese) or {and } (English)
%%%             for the last enumerated Supervisor; Or
%%%             nothing {} for all other supervisors 
%% Optional SecondSuperviser information
% \SupervisorMember{co-orientador}{auxiliar}{do Departamento de Eletrónica, Telecomunicações e Informática}{da Universidade de Aveiro}{e }
% \SupervisorMember{co-orientador}{auxiliar}{do Departamento de Eletrónica, Telecomunicacoes e Informática}{da Universidade de Aveiro}{}

%%% Repeat the above declaration for additional
%%%  Supervisor members as many times as necessary.
%% Additional Supervisor members
% \SupervisoraMember{(co-orientadora)}{associada c/ agregação}{do Departamento de Biologia}{da Universidade de Aveiro}{}
\SupervisorMember{Eurico Pedrosa}{auxiliar}{do Departamento de Eletrónica, Telecomunicacoes e Informática}{da Universidade de Aveiro}{e }

%%% Edit your degree information.
%%%
%%% \Degree{title}{name}
%%%
%%% title: Degree title (Mestre or Doutor)
%%% name : Degree name
%% Degree title and name
\Degree{Mestre}{Robótica e Sistemas Inteligentes}

%%% OPTIONAL: edit grant support texts where
%%%  appropriate. Maximum two.
%%%
%%% \GrantText{grant1}{grant2}
%% Optional Grant support texts
% \GrantText{Texto Apoio financeiro do POCTI no âmbito do III Quadro Comunitário de Apoio.}
%           {Texto Apoio financeiro da FCT e do FSE no âmbito do III Quadro Comunitário de Apoio.}

%%% OPTIONAL: Add a dedication text.
%% Optional Dedication text
\Dedication{}

%%% OPTIONAL: Add an acknowledgement text
%% Optional Acknowledgement text
% Cover page


\Acknowledgements{Esta dissertação consiste no culminar de um percurso académico que se estendeu ao longo de 17 anos. Deste modo, gostaria de começar por agradecer à minha família, especialmente à minha mãe e ao meu pai, por me terem proporcionado esta aventura, bem como por terem celebrado todas as minhas conquistas.

Em segundo lugar, gostaria de agradecer aos meus orientadores, Professor Doutor Bernardo Marques e Professor Doutor Eurico Pedrosa, pelo apoio e orientação na conceção deste trabalho, bem como pelas diversas horas nas quais se dispuseram a ajudar-me, de modo a poder levar este projeto a bom porto. Gostaria também de agradecer ao Doutor João Alves, por me ter transmitido o seu conhecimento e experiência, fundamentais para o desenvolvimento deste trabalho.

Por fim, agradeço também a todos os amigos que fiz ao longo deste percurso, tendo um especial carinho pela “família” que criei em Aveiro, por terem partilhado comigo belos momentos e pela constante presença e apoio nos últimos cinco anos. Especialmente aos “lutadores” que me acompanharam dentro e fora da biblioteca nas últimas semanas: a Matilde, o Ornelas, o Zé, o Mateus, o Canha, o Chico, o Manu, a Catarina, o André, o Silva, o Marcos, o Zé, a Inês, a Carolina, o Pedro, a Casais, a Sara e o Jacinto.

A todos aqueles que se cruzaram comigo e, de algum modo, contribuíram para o meu crescimento pessoal e académico, o meu muito obrigado.}


\AIUse{Reconheço a utilização do ChatGPT (Open AI, https://chat.openai.com/) e do GitHub Copilot (https://github.com/features/copilot) para melhorar a escrita académica, fornecer sugestões de código e assistência no desenvolvimento do software.

\vspace*{5mm}

I acknowledge the use of ChatGPT (https://chat.openai.com/) and GitHub Copilot (https://github.com/features/copilot) for refining academic language and offering code suggestions, aiding in the development of the software.}

%%% OPTIONAL: Acknowledge the use of AI Tools
%\AIUse{I acknowledge the use of [insert AI system(s) and link] to [specific use of generative artificial intelligence or other tasks].
%I acknowledge the use of [software, codes or platforms] to [specific use software, codes or platforms or to other tasks].
%}
%Reconheço o uso de [inserir sistema(s) de IA e respetiva hiperligação] para [indicar utilização específica de IA ou outras tarefas]. Reconheço a utilização de [indicar software, código ou plataforma] para [indicar utilização específica de software, código ou platafoma ou para outras tarefas].
%Exemplo 1: Reconheço a utilização do ChatGPT 3.5 (Open AI, https://chat.openai.com) para resumir as notas iniciais e para rever o rascunho final, a utilização do Office365 (Microsoft, https://www.office.com) para a escrita do texto.
%Exemplo 2: Não foram utilizados no presente trabalho quaisquer conteúdos gerados por tecnologias de IA.


%%% Edit the keywords and abstract in Portuguese
%%% Don't forget to remove the \blindtext[3]
%%%  declaration.
%%% Don't remove the \selectlanguage statement or you will end with incorrect hyphenation
%% Mandatory Keywords and Abstract in Portuguese
\Abstract{palavras-chave}{Indústria 5.0, Colaboração Humano-Robô, Realidade Mista, Gêmeo Digital, Colaboração Remota}
        {resumo}{\selectlanguage{portuguese}A transição para a Indústria 5.0 representa uma mudança significativa cujo objetivo consiste em reintroduzir o elemento humano no ambiente de produção, permitindo uma colaboração mais eficiente entre os operadores e os sistemas inteligentes da Indústria 4.0.  
        A Colaboração Humano-Robô (CHR) proporciona uma maneira eficaz de facilitar a interação entre estes elementos, porém fazê-lo à distância ainda representa um dos maiores desafios desta transição. Face a esta lacuna, surge a Realidade Mista (RM), uma tecnologia revolucionária capaz de a colmatar, ao criar ambientes imersivos onde elementos físicos e virtuais coexistem e interagem em tempo real.
        \\A presente dissertação tem como objetivo responder às novas necessidades da indústria, através do desenvolvimento de uma nova arquitetura que integra a RM com um braço robótico, proporcionando uma colaboração dinâmica e imersiva entre operadores locais e remotos.
        \\O sistema desenvolvido permite ao utilizador local manipular o robô através de uma interface intuitiva composta por elementos visuais e auditivos,  cujo propósito consiste em aumentar a sua imersão e segurança. As zonas de segurança virtuais implementadas pretendem garantir uma interação segura, parando automaticamente o robô caso sejam ultrapassadas. Os utilizadores remotos podem visualizar o cenário de colaboração em tempo real e controlar o robô através da aplicação desenvolvida. Foi adicionada uma câmara ao robô de modo a transmitir a sua perspetiva, permitindo ao utilizador remoto ter uma melhor percepção do cenário de colaboração.
        \\Resumindo, foi desenvolvido um sistema abrangente que permite a integração de novas tecnologias, lançando as bases para melhorias futuras alinhadas com o crescimento e avanço dos cenários de colaboração na indústria. Sendo assim, esta dissertação procura contribuir para ambientes industriais mais eficientes e modernos, tendo em conta os valores centrados no ser humano na Indústria 5.0, promovendo práticas industriais mais eficientes, seguras e colaborativas.}

\Abstract{keywords}{Industry 5.0, Human Robot Collaboration, Mixed Reality, Digital Twin, Remote Collaboration}
        {abstract}{\selectlanguage{english}The transition to Industry 5.0 marks a significant shift toward human-centric manufacturing, emphasizing seamless collaboration between human operators and intelligent machines. A critical challenge in Human-Robot Collaboration (HRC) is facilitating effective remote interaction, enabling humans and robots to cooperate across distance. Mixed-Reality (MR) emerges as a transformative technology capable of bridging this gap, by providing immersive environments where physical and virtual elements coexist and interact in real time.
        \\This dissertation addresses the need for enhanced remote collaboration by developing a novel framework that integrates MR with a robotic arm to enable dynamic and immersive cooperation between on-site and remote human operators.
        \\The developed system allows on-site users to manipulate the robot through an intuitive interface enriched with visual and auditory cues, enhancing user awareness and safety. Virtual safety zones are implemented to ensure secure interaction, automatically halting the robot if breached. Remote users can visualize the collaborative scenario in real time and control the robot via the shared MR application, supported by a live camera feed from the robot's perspective, providing comprehensive situational awareness.
        \\All in all, a comprehensive framework that enables the integration of novel software and technologies is presented, laying the groundwork for future enhancements aligned with societal growth and evolving industrial scenarios. This dissertation aims to contributes to the advancement of modern manufacturing environments by embracing the human-centric values of Industry 5.0, fostering more efficient, safe, and collaborative industrial practices.}


        % The transition to Industry 5.0 marks a significant shift toward human-centric manufacturing, emphasizing seamless collaboration between human operators and intelligent machines. A critical challenge in \ac{HRC} is facilitating effective remote interaction, enabling humans and robots to cooperate across distance. \ac{MR} emerges as a transformative technology capable of bridging this gap, by providing immersive environments where physical and virtual elements coexist and interact in real time.
        % \\This dissertation addresses the need for enhanced remote collaboration in \ac{HRI} by developing a novel framework that integrates \ac{MR} with a robotic arm to enable dynamic and immersive cooperation between on-site and remote human operators.
        % \\The developed system allows on-site users to manipulate the robot through an intuitive interface enriched with visual and auditory cues, enhancing user awareness and safety. Virtual safety zones are implemented to ensure secure interaction, automatically halting the robot if breached. Remote users can visualize the collaborative scenario in real time and control the robot via the shared \ac{MR} application, supported by a live camera feed from the robot's perspective, providing comprehensive situational awareness.
        % \\Despite challenges encountered during development, mainly due to the lack of available documentation and application examples, such as ensuring accurate synchronization between the physical and virtual environments, the resulting framework demonstrates the potential of \ac{MR} in enhancing \ac{HRC}.
        % \\In conclusion, a comprehensive framework that enables the integration of novel software and technologies is presented, laying the groundwork for future enhancements aligned with societal growth and evolving industrial scenarios. This dissertation aims to contributes to the advancement of modern manufacturing environments by embracing the human-centric values of Industry 5.0, fostering more efficient, safe, and collaborative industrial practices.

         
%%% Edit the Jury President information.
%% First Member will be the President of the Jury
\JuriMember{Prof. Doutor José Nuno Panelas Nunes Lau}{professor associado da Universidade de Aveiro}

%%% Repeat the above declaration for additional
%%%  members of the jury as many times as necessary.
%% Additional members of the Jury
\JuriMember{Prof. Doutor Miguel Ângelo Correia de Melo }{professor Auxiliar da Universidade de Trás-Os-Montes e Alto Douro - Escola de Ciências e Tecnologia}
\JuriMember{Doutor Bernardo José Santos Marques}{professor Auxiliar em Regime Laboral da Universidade de Aveiro}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%       DEFINE THESIS COVER FIELDS
%%                  END
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

%%% By default we consider portuguese to be the
%%%  main language (the last one in the usepackage
%%%  declaration). If you're writing your thesis
%%%  in english uncomment the following line.
%%% Set the proper language or you will end with incorrect hyphenation!
\selectlanguage{english}

\MakeCover
\end{document}
